---
title: "Homework 2"
format: html
editor: visual
---

# Homework 2

## Task 1: Basic Vector practice

For this first task, we created vectors for pre-treatment and post-treatment from a medical experiment on blood pressure. We assigned a name to these vectors, and then calculate difference between the respective vectors to determine average difference.

```{r t1-p1}
# Creating Initial Vectors
pre_tr <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post_tr <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

```{r t1-p2}
# Vector and patients IDs
subject_ids <- paste("Subject", 1:20, sep = "_")

names(pre_tr) <- subject_ids
names(post_tr) <- subject_ids
```

```{r t1-p3}
# The change in blood pressure across all patients
diff_bp <- pre_tr - post_tr
diff_bp
```

```{r t1-p4}
# Calculating the average in blood pressure across all patients
avg_diff <- mean(diff_bp)
avg_diff
```

The average of all 20 patients the blood pressure dropped by 17 points.

```{r t1-p5}
# Find subjects with a positive blood pressure change
decreased_bp <- which(diff_bp > 0)
decreased_bp
```

Based on the output, we can see how subjects 9,13, and 17 are not part of the list, indicating that their blood pressure went up after the treatment. Thus, the treatment didn't help them.

```{r t1-p6}
# Return only positive blood pressure changes
positive_diff <- diff_bp[diff_bp > 0]
positive_diff

# Other method used
positive_diff <- diff_bp[decreased_bp]
positive_diff
```

In this case we know that subjects 9, 13 and 17 are the patients whose blood pressure went up after the treatment.

```{r t1-p7}
avg_diff_decreased <- mean(positive_diff)
avg_diff_decreased
```

Comparing the two resulted averages, notice how when we included everyone, the average decreased by 17, however this resulted included 3 patients whose blood pressure did not decrease. After we calculated the average only including the 17 patients who blood pressure improved, the average decreases about 20.65 points. This results is more accurate to identify the effectiveness of the treatment.

## Task 2: Basic Data Frame Practice

For this second task, we built a data frame the previous vectors, filtered for negative differences in blood pressure, added a normal range column which evaluated if blood pressured is less than 120, and printed this data frame using the `kable()` function.

```{r t2-p1}
# Create the blood pressure data frame
bp_df <- data.frame(
  patient = subject_ids,
  pre_bp = unname(pre_bp),
  post_bp = unname(post_bp),
  diff_bp = unname(diff_bp)
)

bp_df

# Other method
patient <- 1:20

bp_df <- data.frame(patient = patient,
                    pre_bp  = pre_tr,
                    post_bp = post_tr,
                    diff_bp,
                    row.names = NULL)
bp_df
```

```{r t2-p2}
# Returning only rows where the diff_bp column is negative
subset(bp_df, diff_bp < 0)
# Other method
negative_diff <- bp_df[bp_df$diff_bp < 0, ]
negative_diff
```

```{r t2-p3}
# Adding a new column named normal_range to the data frame that is TRUE if the post_bp is less than 120
bp_df$post_under_120 <- bp_df$post_bp < 120
bp_df
```

```{r t2-p4}
# Print the data frame with knitr
knitr::kable(bp_df)
```

## Task 3: List Practice

For third section, we built the placebo data frame (same as Task 2), then we combined both data frames into a named list, and demonstrated a different ways to access elements from that list.

```{r t3-p1}
# Pre-treatment blood pressure measurements for placebo subjects
placebo_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
# Post-treatment blood pressure measurements for placebo subjects
placebo_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

# Calculate the change in blood pressure
placebo_diff <- placebo_pre - placebo_post

# Create the placebo data frame
placebo_df <- data.frame(
  patient = subject_ids,
  pre_bp = placebo_pre,
  post_bp = placebo_post,
  diff_bp = placebo_diff
)



# Other method
subject_ids  <- 1:10

placebo_df <- data.frame(
  patient = subject_ids,
  pre_bp = placebo_pre,
  post_bp = placebo_post,
  diff_bp = placebo_diff,
  row.names = NULL
)

# Other method 2
placebo_df <- data.frame(patient=subject_ids,
                        pre_bp=placebo_pre,
                        post_bp=placebo_post,
                        diff_bp=placebo_diff,
                        normal_range=Post_placebo < 120,
                        row.names = NULL)



# Add the same logical column used in Task 2
placebo_df$post_under_120 <- placebo_df$post_bp < 120
placebo_df
```

```{r t3-p2}
# Combine the treatment and placebo data frames into a list
bp_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)

bp_list

# Other method
experiment_list <- list(treatment = bp_df, placebo = placebo_df)
experiment_list

```

```{r t3-p3}
# Access by numeric position
bp_list[[1]]

# Access by element name using double brackets
bp_list[["treatment"]]

# Access by element name using $
bp_list$treatment

# Other method using the experiemnt_list
experiment_list[[1]]
experiment_list[["treatment"]]
experiment_list$treatment
```

```{r t3-p4}
# Access the placebo pre_bp column
bp_list$placebo$pre_bp
experiment_list$placebo$pre_bp
```

## Task 4: Control Flow Practice

For this last section, we added a status column to each data frame inside the list, then used a for loop with if/else logic to classify each patient as optimal, borderline, or high based on their post treatment/placebo blood pressure values.

```{r t4-p1}
# Create empty status columns in the data frames stored in the list
bp_list$treatment$status <- character(nrow(bp_list$treatment))
bp_list$placebo$status <- character(nrow(bp_list$placebo))

# Other method
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```

```{r t4-p2}
for (i in 1:nrow(experiment_list$treatment)) {
  bp_value <- experiment_list$treatment$post_bp[i]
  if (bp_value <= 120) {
    experiment_list$treatment$status[i] <- "optimal"
  } else if (bp_value > 120 & bp_value <= 130) {
    experiment_list$treatment$status[i] <- "borderline"
  } else {
    experiment_list$treatment$status[i] <- "high"
  }
}

experiment_list$treatment


# Other method

for (i in 1:20) {
  
  x <- bp_list$treatment$post_bp[i]
  
  if (x <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  } 
  
  else if (x > 130) {
    bp_list$treatment$status[i] <- "high"
  } 
  
  else {
    bp_list$treatment$status[i] <- "borderline"
  }
  
}
```

```{r t4-p3}
for (i in 1:nrow(experiment_list$placebo)) {
  bp_value <- experiment_list$placebo$post_bp[i]
  if (bp_value <= 120) {
    experiment_list$placebo$status[i] <- "optimal"
  } else if (bp_value > 120 & bp_value <= 130) {
    experiment_list$placebo$status[i] <- "borderline"
  } else {
    experiment_list$placebo$status[i] <- "high"
  }
}

experiment_list$placebo

# Other method

for (i in 1:10) {
  
  x <- bp_list$placebo$post_bp[i]
  
  if (x <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  } 
  
  else if (x > 130) {
    bp_list$placebo$status[i] <- "high"
  } 
  
  else {
    bp_list$placebo$status[i] <- "borderline"
  }
  
}
```
