---
title: "Homework 3"
author: "Javi B"
editor: visual
---

## Task 1: Conceptual Questions

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

2.  What are the major benefits of using R projects?

3.  What is Git and what is GitHub?

4.  What are the two main differences between a tibble and a data.frame?

## Task 2: Reading Delimited Data

### Glass Data

1.  Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name).

    ```{r t2-p1}
    library(tidyverse)
    glass_data <- read.csv(
      "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
      col.names = c("ID", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")
    )
    glass_data
    ```

### Yeast Data

2.  Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name)

    ```{r t2-p2}
    yeast_data <- read_table(
      "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
      col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
    )

    yeast_data
    ```

## Task 3: Combining Excel and Delimited Data

3.  There is an excel version of the white wine data set available at <https://www4.stat.ncsu.edu/~online/> datasets/white-wine.xlsx.

    - Download this file

    - Place it in a folder you know (such as your working directory for your project, the folder your .qmd file lives in, a data folder you’ve created in your repo, etc.)

    - Import the data from the first sheet using the readxl package

    - Print out the tibble (just call the object name)

    ```{r t3-p3}
    library(readxl)

    white_wine <- read_excel(
      "HW3/data/white-wine.xlsx",
      sheet = 1
    )
    white_wine
    ```

4.  When you print the data set out to the console, you may notice that some of the variable names are surrounded by backticks. This is because they are non-standard (they include a space in them). We can rename them in a number of ways. We’ll do it by reading in the variable names from the 2nd sheet of the same file.

    - Read in the data from the 2nd sheet. This should return a data frame with one column containing alternative versions of the variable names.

    - Grab that column and overwrite the current column names (`colnames()`) of your white wine tibble. Note that `colnames(object)` allows for ‘left hand side assignment’. That is, you can call it on the left and store an object in it to overwrite it. Similar to `names(object) <- c(...)`

```{r t3-p4}
colnames(white_wine) <- read_excel("HW3/data/white-wine.xlsx", sheet = 2) |>
  pull(1)
```

5.  Lastly, add a column to this data set to indicate the wines are white. That is, add a column that has values of ‘white’ for every observation. You can do this via the \$ operator or, if you know it already, the mutate() function from `dplyr`.

    ```{r t3-p5}
    white_wine <- white_wine |>
      mutate(type = "white")
    ```

6.  There is a semi-colon delimited version of the red wine data set available at <https://www4.stat.ncsu.edu/~online/> datasets/red-wine.csv.

    - Read this in using the `readr` package. Be careful that the columns are read in as the correct type!
    - You should replace the variable names as done above
    - You should append a column denoting the type as “red”

    ```{r t3-p6}
    library(readr)

    red_wine <- read_delim(
      "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
      delim = ";"
    )
    colnames(red_wine) <- read_excel("HW3/data/white-wine.xlsx", sheet = 2) |> 
      pull(1)
    red_wine <- red_wine |> 
      mutate(type = "red")
    ```

7.  Combine these two data sets into one data set. They both have the exact same columns so this is an easy append task!

    - Use the `dplyr::bind_rows()` function (see the help) to create one tibble containing all of the wine data.

    ```{r t3-p7}
    wine_data <- dplyr::bind_rows(white_wine, red_wine)
    wine_data
    ```