---
title: "Homework 3"
author: "Javi B"
editor: visual
---

## Task 1: Conceptual Questions

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

2.  What are the major benefits of using R projects?

3.  What is Git and what is GitHub?

4.  What are the two main differences between a tibble and a data.frame?

## Task 2: Reading Delimited Data

### Glass Data

1.  Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name).

    ```{r}
    library(tidyverse)
    glass_data <- read.csv(
      "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
      col.names = c("ID", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type")
    )
    glass_data
    ```

### Yeast Data

2.  Read this data into R directly from the URL using functions from the tidyverse. Notice that the data doesn’t include column names - add those (in a manner of your choosing). Print out the tibble (just call the object name)

    ```{r}
    yeast_data <- read_table(
      "https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
      col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
    )

    yeast_data
    ```